### 2019-05-13 paper/response_to_referee.txt
more referee items

Dear Editor:

Enclosed please find the author response to the referee report on
AAS14152 "Spectrophotometric parallaxes with linear models: Accurate
distances for luminous red-giant stars" by Hogg, Eilers, & Rix.

Apologies for the tardiness of this response. The referee has
definitely made great suggestions and the paper is improved for us
having addressed them. Our response to the referee is below the line.

David W Hogg, for the authors

-----

In what follows, the referee comments are indented, and our responses
are not. We have made changes to the paper in response to every point;
these comments were helpful and valuable.

    The paper presents a method to estimate spectrophotometric parallaxes
    based only on input data, without adoption of physically motivated
    models and priors. The paper is written in an way that is strongly
    focused on the method itself, but leaves several complementary aspects
    that are certainly important to assess the reliability of the inferred
    parallaxes and, most importantly, distances. I recommend its
    publication only after such major issues are addressed.

    Major issues

    1 - The data-driven method described in the paper details out how
    spectrophotometric parallaxes are inferred. But the paper does not
    give the reader a clue about how spectrophotometric distances are
    estimated. The authors make it clear in Section 2 that they never
    invert *observed* parallaxes to estimate distances in their
    method. But are the spectrophotometric distances they infer (shown in
    Figures 5 and 6) just the inverse of the *inferred* parallaxes? If so,
    this should be made explicit in the text, along with some discussion
    about the problems (or lack of) that may arise from this
    inversion. Even though the spectrophotometric parallaxes themselves
    are always positive, a reader would like to be convinced that the
    distances shown in Figures 5 and 6 do not suffer from biasses caused
    by the asymmetric nature of the distances distribution that results
    from a symmetric error distribution in parallaxes (as stated by the
    likelihood assumption of the method). How do these biasses simply go
    away when the method's w(sp) +/- sigma(sp) are used? In particular, a
    negative parallax for many stars may be consistent with their w(sp)
    +/- sigma(sp).

These issues are subtle! We are assuming that the GAIA-MEASURED
PARALLAXES have noise that is symmetric in parallax space. In our
data-driven model, we use a linear combination of features to predict
or estimate a LOG PARALLAX, so if anything we are claiming that our
estimates (not the data, but our estimates based on the data) are
symmetric in log parallax, not parallax. But we aren't even really
claiming that. The issues are subtle. However, this does deserve some
discussion and we have added that to the captions of Figure XXX and
YYY (formerly Figure 5 and 6), and in the last paragraph of the
discussion, for good emphasis [HOGG DO THIS AT "PLACEHOLDER A"].  The
short answer is that we just exponentiated the negative of our
log-parallax estimates to make those figures.

There is an additional question of whether we are somehow biasing
things when we presume that the parallax must be positive? Our
position is that we are biasing things by presuming that the LOG
PARALLAX can be predicted as a linear combination of features. Of
course any other assumption would also introduce biases! There is no
way to fit that would be unbiased here, because by construction, our
model is not physically accurate. We added words about this into the
linearity assumption in the assumption list in Section 2.

    2 - The paper shows limited validation tests and comparisons to other
    work. There are several papers that provide spectrophotometric
    distances based on physical models for APOGEE stars (Queiroz et al
    2018, MNRAS, 476, 2556; Schultheis et al 2014, AJ, 148, 24 and
    others). The authors should compare their distances to those authors
    and discuss the results of this comparison. Moreover, asteroseismic
    distance estimates are also available for a fraction of the APOGEE
    sample of giants in common with the CoRoT and Kepler missions. This is
    an additional and important validation test.

This is a great point. We obtained these two data sets (and the
Queiroz sample required some emails!) and we made comparisons. We
added a short paragraph about the results of these comparisons in
Section XXX. We did not feel that the comparisons merited supporting
Figures, but we do have such figures if the referee disagrees. HOGG DO THIS

During the long period in which we haven't responded to this report,
there was also a paper out by Leung & Bovy. They compare their results
to ours, and we also mention that in the newly added paragraph. HOGG DO THIS

    3 - The authors apply colour cuts to eliminate outliers in the APOGEE
    data, eqs 6 and 7. They also apply a cut in parallax error, eq. 8.,
    along with other quality criteria, to define the training set. So, the
    training set parallax uncertainty distribution is truncated. How do
    these cuts affect the method and its assumptions? In particular, it is
    important to reconcile these selection criteria for the training set
    with the statements presented in the Introduction (page 3):
    "... regressions can be strongly biased if we have a bad selection
    function or data censoring. This means that we cannot apply cuts to
    the Gaia astrometry or signal-to-noise unless those cuts are correctly
    accounted for in our generative model for the data sample. Below, we
    will make no such cuts."  It is in fact clear from Figure 1 that the
    training set is well under represented of low log g stars with respect
    to the parent sample. This already jeopardizes the representativeness
    assumption somewhat.  One useful assessment of the effects of data
    censoring would be to show the spectral S/N or parallax distributions
    of the training data in comparison to that of the entire sample that
    the method is applied to.

This point is important. The Gaia parallax uncertainties are based on
a spacecraft model that is independent of the measurements themselves.
Therefore, cuts on parallax uncertainties are not cuts on parallax and
thus are not subject to the strong censure that we give to such cuts!
But the referee is correct that we have been very unclear about this
and we have attempted to fix this in the relevant paragraph towards the
end of Section 2. This has improved the paper.

The referee is completely correct that the detailed distribution of
stars in parallax, SNR, and log g is not identical in the training set
and the parent sample.  We don't have any reason to believe that this
biases our results. However, we do now show these distribution
differences in a newly added figure, Figure XXX. Certainly these
differences mean that we probably have uncertainties that, in detail,
vary with log g. HOGG DO THIS

    4 - Finally, it is important to provide the reader with some
    understanding of how in practice the data are driving the resulting
    parallaxes. This may be achieved by discussing the behaviour of the
    linear coefficients theta. How do their values, as estimated by
    expr. (3), vary from training set A and to training set B? What
    spectral ranges are most relevant and survive the L1 regularization
    process? In other words, the authors could provide a discussion
    similar to the one around Figure 5 of Ness et al did on the Cannon,
    but this time comparing sets A and B across the spectrum or over the
    smaller photometric subspace.

Absolutely! Actually we had a section about this but dropped it just
before submission. One of the reasons we dropped it is because the
story is somewhat complicated. But you are right that this must be
there, and we have added it back in, at the end of Section YYY. The
short story (which you can read in the updated text) is that the
photometric part of the model is interpretable and stable, but the
spectroscopic part is less stable in its internals (while remaining
very stable in its predictions in the data space). We don't fully
understand this, but it connects to known issues in machine learning
more generally, which we point out in the text. HOGG DO THIS

    Other issues:

    Introduction, 1st par.: could the authors say a few words about the
    importance of making "precise kinematic and element-abundance maps of
    the Milky Way disk out to large heliocentric radii..."?

Good point! The paper now opens with the big question and flows to the
spectroscopy and precision. This change affects the first paragraph of
the paper and changes how it starts.

    Section 2, 1st par. after the list of assumptions. The statement "We
    never cut on parallax or parallax signal-to-noise." is repeated.

Yes, yes it is. Fixed and removed.

    Section 4: "The hope of the model assumptions is that the Training Set
    will contain sufficient dust variation that the model will naturally
    learn the dust corrections and extrapolate acceptably.” Is this hope
    checked in any way?

Hahaha no. Okay fine. We now comment on this point in Section YYY
where we deliver the interpretation of the photometric part of the
model. HOGG DO THIS.

    Section 4: The parallax offset w_0 is a contentious issue and it is
    important to quantify its effect on the final spectrophotometric
    parallaxes.

The effect is very weak; we don't find that the change from 0.029 to 0.048
makes much difference to our results except at very small parallaxes. And
our 0.048 is not preferred over 0.029 at large confidence. The main reason
it has little effect is that it effectively moves the training data only a
tiny fraction of a sigma. We added two sentences to this effect in the relevant
place.

    The caption of Figure 2 is wrong: "...shows Gaia astrometric
    parallaxes w(a)n on the HORIZONTAL axis and our spectrophotometric
    parallaxes w(sp) on the VERTICAL axis."

Whoops! Fixed.

    The panels and labels of Figure 3 are too small. Impossible to read
    the axes.

Fixed. This is a much better figure now.

    Make an analogous of Figure 2 and 3 for all the stars with available
    parallaxes, regardless of whether criteria 8 and 9 are valid or not

Done. HOGG: DO WE PUT THIS INTO THE PAPER?

    Section 5, 2nd paragraph: Is the parameter Lambda (capital letter) the
    same as lambda (lower case initial)? If so, what is meant by the value
    of the "lambda vector" which the authors mention in the same
    paragraph? Please clarify

Whoops. That's just a typo. Fixed. We adjusted that whole paragraph,
which was word salad, into something that is not wrong and also
clearer.

    Typos:

    4th line after eq 5: "... is, we ASSUME THAT C_m is a diagonal matrix
    with the..." (instead of "... we make C_m is ...")

fixed

    Section 4, 2nd par., 4th line: please eliminate repeated "the" in "In
    detail, we obtain the the APOGEE..."

fixed

    Section 4, 4th par., 5th line: "...3.6 and 4.5 μmPERIOD In detail..."

fixed

    Section 4, 5th par., 2nd line: again, eliminate extra "the" in "...to
    the the 2MASS IDs"

fixed

    Section 4, 4th line below eq 10: eliminate extra "or" in "...cut ever
    on parallax or or the ratio..."

fixed

    Comments of the AAS Data and Software reviewer:

    - We recommend reproducibility material such as that hosted at the
      aceilers/spectroscopic parallax repository be archived to a
      repository, with the repository DOI added adjacent to the github URL
      in the final article. [This request is also true of submission
      AAS14147 for similar reasons].

Fixed the zenodo URL to a doi.org URL.

We will do a final URL for the repo on acceptance; we can't do it
before, because git hash! Happy to work to ensure reproducibility!

    - Do the authors wish to see the full Table 1 hosted with the Journal?
      We'd gladly process the final table into the final article, which
      would speed ingest of that data into CDS/Vizier. Regardless the
      Zenodo URL in footnote # should be converted back to its DOI form.

Yes we would like this. Let's discuss at acceptance.

In addition to the above we made a few other small changes:

- We added a reference to the Leung paper that followed ours in the
relevant part of the discussion (where we discuss machine-learning
alternatives).