### 2019-07-31 paper/response_to_referee2.txt
response2

Dear Editor:

Enclosed please find the author response to the referee report on
AAS14152R1 "Spectrophotometric parallaxes with linear models: Accurate
distances for luminous red-giant stars" by Hogg, Eilers, & Rix.

Apologies for the tardiness of this response. Our response to the
referee is below the line.

David W Hogg, for the authors

-----

In what follows, the referee comments are indented, and our responses
are not. These comments improved the paper; we appreciate the close
read!

   page 3: "... hope to model are smoothPERIOD In that case, much of
   the information in the data can ..."

Done.

   page 11: Section 5, 2nd paragraph: please use the same verbal tense
   in the sentences.

We disagree here. The first two sentences say what we did, and the last
says what we find, having done that.

   page 12, 2nd par: "... metallicity, age and extinction,..." (I
   suggest to replace the first "and" by a comma)

Done.

   page 13, par. that compares to other distance estimates: "the
   comparison with the more recent sample (Queiroz et al. 2018) IS
   better:"

Done.

   page 13, par. that compares to other distance estimates: remove ")"
   after "at distances < 8 kpc"

Done.

   page 15, section 6, 3rd par.: "...our choice of features is very
   rigidPERIOD In principle..."

Done.

   In fact, in this last section there several other examples where
   colon was used where a period would better separate the sentences,
   in my humble opinion!

Hahaha! Understood. Removed some.

   page 15, last par.: Perhaps replace "non-worse", which sounds
   weird, by "similar quality"?

Done.

   top of page 16: "...are manifold, howeverPERIOD We would have had..."

Done.

   page 16, end of 3rd par.: "...that we did not OBTAIN as much insight.."

Done.

   Section 6: perhaps the authors could shorten a bit the comparison
   with more complex models based on machine learning and the
   discussion on the advantages and disadvantages of the linear model
   adopted.

Shortened. That improves this paragraph.

   In this last section, there are also a few sentences that are
   repeated from previous sections, such as: "This framework ensures
   that the Gaia parallax ... in generating the spectrophotometric
   parallax ... That means that the parallax estimates generated here
   and given in Table 1 are statistically independent of the Gaia
   parallaxes" This sentence is basically repeated from the end of
   Section 3.

Rephrased these; we do want to emphasize these points by repetition,
so we didn't delete them, just rephrased them to make it clear that
it is a repetition for emphasis.